---
title: "R For Data Sciences - Exercises"
author: "Hans Franke"
date: "September 29, 2020"
output: rmarkdown::github_document
---

```{r}
library(tidyverse)
library(ISLR)
```

# Chapter 3: Data Visualisation

## 3.2.4 Exercises

Run ggplot(data = mpg). What do you see?  

*Answer:* Nothing, because there is no axis, on aes1 aditional must define GEOm to show proper graph.

```{r}
  # corrected code
  #View(mpg)

  ggplot(data = mpg, aes(x = manufacturer, y = hwy))+
  geom_point()
```


How many rows are in mpg? How many columns?

```{r}
  # 234 rows and 11 columns
  str(mpg)
```


What does the drv variable describe? Read the help for ?mpg to find out.

```{r}
  ?mpg

  # DRV = the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd
```

Make a scatterplot of hwy vs cyl.

```{r}
  ggplot(mpg, aes(hwy,cyl))+
  geom_point()+
  theme_minimal()
```


What happens if you make a scatterplot of class vs drv? Why is the plot not useful?
```{r}
#Both categorical variables, scatterplot usually shows some relations between variables, but to get better results with one continuos.
  ggplot(mpg, aes(class,drv))+
  geom_point()+
  theme_minimal()
```


## 3.3.1 Exercises

What’s gone wrong with this code? Why are the points not blue?



```{r}
#Color inside the AES, determine CLASS of points, usually is used by other variable to describe or add informations
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```


Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

Answer:
- Continuos: displ, year, cty, hwy
- Categorical: manufacturer, model, trans, drv, fl, class
```{r}
#Look for variables NON-Numeric => categorical
str(mpg)
```


Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
```{r}
#shape => dosent accept continuos 
# size and color => accpect but i will not recommend for continuos variables with many values!
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cty, size = year))
```


What happens if you map the same variable to multiple aesthetics?

```{r}
#it will be shown the same information.... lot of unecessary visual with no value added!
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class, size = class, shape = class))
```


What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

```{r}
#stroke = define the size of the points...
  ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy, stroke = 2 , color = class, shape = class))
```


What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.

```{r}
  #Split variable COLOR to true or false with logical conditional 
  ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy, stroke = 2 , colour = displ < 5, shape = class == "minivan"))
```



## 3.5.1 Exercises
What happens if you facet on a continuous variable?

What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
What plots does the following code make? What does . do?

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
Take the first faceted plot in this section:

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?

When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

## 3.6.1 Exercises
What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

What does the se argument to geom_smooth() do?

Will these two graphs look different? Why/why not?


ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
  
  
## 3.7.1 Exercises

What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

What does geom_col() do? How is it different to geom_bar()?

Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?

What variables does stat_smooth() compute? What parameters control its behaviour?

In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?


ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
  
  
## 3.8.1 Exercises
What is the problem with this plot? How could you improve it?

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()


What parameters to geom_jitter() control the amount of jittering?

Compare and contrast geom_jitter() with geom_count().

What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.

## 3.9.1 Exercises

Turn a stacked bar chart into a pie chart using coord_polar().

What does labs() do? Read the documentation.

What’s the difference between coord_quickmap() and coord_map()?

What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?


ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
  
  K