---
title: "R For Data Sciences - Exercises"
subtitle: "Chapter 7 - EDA"
author: "Hans Franke"
date: "September 30, 2020"
output: rmarkdown::github_document
---


```{r}
library("nycflights13")
library("tidyverse")
```


# 7.3 Variation

## 7.3.4 Exercises
- Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.
```{r}
# x, y and Z are continuos, so we should use a histogram!
 str(diamonds)
```
```{r}
#Some summary statistics

#1) how u can have some diamonds with 0 on their dimensions?
#2) probably the max values from y and z are wront too, they are so high compare with mean/median
#3) Looking at plots (see below) we notice there are some outliers, looking for the space in the x-axis
#4) looking for help ?diamonds, we notice that x = lenght, y = width and z = deph. 
summary(select(diamonds, x, y, z))

```

```{r}
  
  ggplot(diamonds, aes(x))+
  geom_histogram(fill = "dark blue") +
  theme_minimal()
```
```{r}
  
  ggplot(diamonds, aes(y))+
  geom_histogram(fill = "dark green") +
  theme_minimal()
```

```{r}
  
  ggplot(diamonds, aes(z))+
  geom_histogram(fill = "dark red") +
  theme_minimal()
```


- Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)

```{r}
#there is no diamong with price in that "hole"
ggplot(diamonds, aes(price))+
  geom_histogram(binwidth = 10, center = 0)+
  coord_cartesian(ylim= c(0,1000), xlim=c(0,2000))
```


- How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?

```{r}
#there is much more diamonds with carat =1
diamonds %>%
  filter(carat>0.98 & carat <1.01) %>%
ggplot(aes(carat))+
  geom_histogram()
```
```{r}
#showing the values
diamonds %>%
  filter(carat >= 0.99, carat <= 1) %>%
  count(carat)
```



- Compare and contrast coord_cartesian() vs xlim() or ylim() when zooming in on a histogram. What happens if you leave binwidth unset? What happens if you try and zoom so only half a bar shows?

```{r}
#the coord_cartesian works after the bins were evaluated
ggplot(diamonds) +
  geom_histogram(mapping = aes(x = price)) +
  coord_cartesian(xlim = c(100, 5000), ylim = c(0, 3000))

```


```{r}
#x_lim and y+lim cuts from the plot values outside the range values! So the plot can be different
ggplot(diamonds) +
  geom_histogram(mapping = aes(x = price)) +
  xlim(100, 5000) +
  ylim(0, 3000)
```

# 7.4: Missing Values

## Question: 7.4.1

What happens to missing values in a histogram? What happens to missing values in a bar chart? Why is there a difference?


*Answer:* 
- Histogram => missing values are when the number of observations in each bin are calculated.
- Bar_chart => n/a values are a another category

What does na.rm = TRUE do in mean() and sum()?

*Answer: This option removes NA values from the vector prior to calculating the mean and sum*

# 7.5: Covariation

## 7.5.1.1 Exercises
Use what you’ve learned to improve the visualisation of the departure times of cancelled vs. non-cancelled flights.

```{r}
flights %>% 
  mutate(
    cancelled = is.na(dep_time),
    sch_dep = sched_dep_time %/% 100,
    sch_dep_min = sched_dep_time %% 100,
    sche = sch_dep + sch_dep_min / 60
    
  ) %>%
  ggplot(aes(x = sche, y = cancelled)) + 
  geom_boxplot()
```


What variable in the diamonds dataset is most important for predicting the price of a diamond? How is that variable correlated with cut? Why does the combination of those two relationships lead to lower quality diamonds being more expensive?

```{r}
# Variables: carat, cut, color, cut
head(diamonds)
```
```{r}
#Carat vs Price -> 2 continuos so we should use scatterplot
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point()
```
```{r}
#Carat vs Price but with boxplot (One continuos cutted with bins) => Positive correlation, as increase in carat increase the price
ggplot(diamonds, aes(x=carat, y = price))+
  geom_boxplot(aes(group=cut_width(carat,0.2)))

```
```{r}
#Cut vs Price, negative correlation (1 continuos x 1 categorical)
ggplot(diamonds, aes(x=cut, y=price))+
  geom_boxplot()
```

```{r}
#Color vs Price, negative correlation (1 continuos x 1 categorical) => this is dificult to visualise but there is a very light negative correlation (color J is the worst)
ggplot(diamonds, aes(x=color, y=price))+
  geom_boxplot()+
  theme_minimal()
```
```{r}
#Clairty vs Price, negative correlation (1 continuos x 1 categorical) => this is dificult to visualise but there is a very light negative correlation (color IF is the worst)
ggplot(diamonds, aes(x=clarity, y=price))+
  geom_boxplot()+
  theme_minimal()
```

## 7.5.2 Two Categorical Variables

```{r}
diamonds %>%
  count(color, cut) %>%
  group_by(color) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(mapping = aes(x = color, y = cut)) +
  geom_tile(mapping = aes(fill = prop))
```


## 7.5.3 Two continuous variables

```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_hex() +
  facet_wrap(~cut, ncol = 1)
```

